---
title: "Analisis ENCOVI 2023"
output: html_notebook
---

# Load packages and create functions

```{r}
library(readr)
library(readxl)
library(zoo)
library(lubridate)
library(dplyr)
library(tidyr)
library(haven)
library(survey)
library(writexl)

```

Function
```{r}
#Function to read excel files
read_excels <- function(folder_path) {
  # Get all the Excel file names in the folder
  excel_files <- list.files(folder_path, pattern = "\\.xlsx$", full.names = TRUE)
  
  # Initialize an empty list to store the data
  data_list <- list()
  i=1
  
  # Loop over each Excel file and read it
  for (file in excel_files) {
    df_name <- paste0('df_', i)
    # Read the Excel file and store it in the list
    data_list[[df_name]] <- read_excel(file)
    i=i+1
  }
  
  # Return the list of data frames
  return(data_list)
}

#Function to read .sav files SPSS
read_savs <- function(folder_path) {
  # Get all the Excel file names in the folder
  sav_files <- list.files(folder_path, pattern = "\\.sav$", full.names = TRUE)
  
  # Initialize an empty list to store the data
  data_list <- list()
  i=1
  
  # Loop over each Excel file and read it
  for (file in sav_files) {
    df_name <- paste0('df_', i)
    # Read the Excel file and store it in the list
    data_list[[df_name]] <- read_sav(file)
    i=i+1
  }
  
  # Return the list of data frames
  return(data_list)
}
```



# Load data

## Load dictionaries
```{r}
path <- "/Users/fabianjp/Documents/Work/ASIES/Desarrollo local/Data/Encovi 2023/Data/SAV"

survey_dfs <- read_savs(path)
```

## Load sav files
```{r}
path <- "/Users/fabianjp/Documents/Work/ASIES/Desarrollo local/Data/Encovi 2023/Diccionario"

dictionaries <- read_excels(path)
```


```{r}
i=1
for (df in survey_dfs){
  print(paste0(i, ' length of ', (nrow(df) * length(df))/1000000))
  i=i+1
}

View(survey_dfs$df_31)
```

df_31 has informatino about the person and has several important characteristics that we can use

Select main tables
```{r}
personas <- survey_dfs$df_31
```


```{r}
# Get a list of variable names
variable_names <- names(personas)

# Get the variable labels
variable_labels <- sapply(personas, function(x) attr(x, "label"))

# Combine them into a data frame for easy viewing
variable_info <- data.frame(variable_names, variable_labels, stringsAsFactors = FALSE)

# Print the first few rows to explore
head(variable_info)

```


# Checks and survey analysis functions

## Create variables

```{r}
personas <- personas %>%
  mutate(edad_cat_1 = case_when(
    PPA03 <= 17 ~ '0-17',
    PPA03 <= 59 ~ '18-59',
    PPA03 > 59 ~ '60 o mas',
    TRUE ~ NA_character_  # Correct syntax for default missing value
  ) %>% factor(levels = c('0-17', '18-59', '60 o mas')))

personas$sexo <- as_factor(personas$PPA02)
personas$departamento <- as_factor(personas$DEPTO)
personas$pobres <- as_factor(personas$POBREZA)
personas$educacion <- as_factor(personas$P06B26A)

```

### Subsets
```{r}
adultos <- subset(personas, PPA03 >= 18)

```


## Checks

Verificar los pesos y si los resultados cuadran con los presentados por el INE en su presentación

Pesos
```{r}
# Create a survey design object

update_survey_design <- function() {
  # Example: Assuming you have a data frame `mydata` with a weight column `weight_var`
survey_design <- svydesign(ids = ~1,           # No clustering (use clustering variable if applicable)
                           data = personas,      # Your data frame
                           weights = ~FACTOR)  # Column with weights

survey_design_adultos <- svydesign(ids = ~1,           # No clustering (use clustering variable if applicable)
                           data = adultos,      # Your data frame
                           weights = ~FACTOR)  # Column with weights
}

```

Variables descriptivas de la población
```{r}
#sexo
svymean(~factor(PPA02), design = survey_design) #proporcion por sexo
svymean(~sexo, design = survey_design) #proporcion por sexo

#edad
svyquantile(~PPA03, design = survey_design, c(0.5)) #edad media
svymean(~factor(edad_cat_1), design = survey_design) #proporcion por edad

#poblacion urbana-rural
svymean(~factor(AREA), design = survey_design) #proporcion por urbano-rural

#educacion
svymean(~educacion, design = survey_design_adultos) #proporcion por educacion

#pobreza
svymean(~factor(POBREZA), design = survey_design) #proporcion por pobreza

#departamento
#svymean(~departamento, design = survey_design)
svytotal(~departamento, design = survey_design) #poblacion por departamento

```

Contingency tables
```{r}
#Pobreza 
svyby(~sexo, ~POBREZA, svymean, design = survey_design) #por sexo
svyby(~factor(POBREZA), ~sexo, svymean, design = survey_design) #por sexo
svyby(~pobres, ~departamento, svymean, design = survey_design) #por departamento


#Edad
svyby(~PPA03, ~departamento, svyquantile, design = survey_design, quantiles = 0.5, ci = TRUE)

#Educacion
svyby(~educacion, ~edad_cat_1, svymean, design = survey_design) #por sexo


```



# Variables de desarrollo

Personas
- Educacion
- Edad
- Empleo

Externalidades
- 

## Variables demograficas para el analisis
```{r}
#demos
personas$sexo <- as_factor(personas$PPA02)
personas$departamento <- as_factor(personas$DEPTO)
personas$pobres <- as_factor(personas$POBREZA)
personas$educacion <- as_factor(personas$P06B26A)
demos <- c('sexo', 'educacion')
```

## Variables de desarrollo
Pobreza: utiliza tres escalas de pobreza basados en los umbrales del INE
```{r}
personas$pobreza <- as_factor(personas$POBREZA)
```

Participacion ciudadana
Importante para el analisis comunitario
```{r}
#Participacion ciudadana
#La persona participa en alguna asociacion de cualquier tipo
personas$participacion_ciudadana <- as_factor(personas$P03B01)

#participacion de enfasis comunitario economico -  la personas solo selecciona la principal
personas <- personas %>% mutate(participacion_comunitaria_economica = case_when(
  P03B03 %in% c(1,8,9,10,13,15,21,22,23,28) ~ 1,
  is.na(P03B03) ~ NA_real_,
  TRUE ~ 0
))

update_survey_design()
variables_des <- c('pobreza', 'participacion_comunitaria_economica')
```


```{r}
df_merge <- svyby(~pobres, ~departamento, svymean, design = survey_design) %>% 
  mutate(pobreza_general = pobres0 + pobres1 + pobres2,
         pobreza_extrema = pobres0 + pobres1) %>% select(departamento, pobreza_general, pobreza_extrema)

df_merge <- df_merge %>% left_join(
  svyby(~participacion_ciudadana, ~departamento, svymean, design = survey_design, na.rm = TRUE) %>%
    rename(participacion_ciudadana = participacion_ciudadanaSi) %>%
    select(departamento, participacion_ciudadana)
)
df_merge

```


```{r}
#Save excel to merge
getwd()
write.csv(df_merge, "/Users/fabianjp/Documents/Work/ASIES/Desarrollo local/Github/R/Desarrollo_local/Data/data_encovi.csv", row.names = FALSE)
```


```{r}
variables_des <- c('pobreza', '')
```



Podemos entender un poco sobre las fuentes y recursos naturales presentes en las localidades?
